---
title: "Computational Musicology"
author: "Lieve"
date: "2/23/2022"
output:
  flexdashboard::flex_dashboard:
    orientation: columns
    storyboard: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(flexdashboard)
library(tidyverse)
library(plotly)
library(spotifyr)
library(ggplot2)

```

# Chromagrams

# Introduction

hallo

## Column 2 {data-width = 650}

### Corpus

Sitting in a cafe or restaurant, music has a big influence on the atmosphere. Working in a cafe myself, I use a lot of spotify playlists. During the day it will be in the trend of 'Coffee House', in the evening more jazz and in the late hours a rock party. I want to find out what aspects of this music makes them more enjoyable on different parts of the day. For this I will use the playlists we play at my work and compare these with each other. This will start with a comparison of genre, which will be very different. The first thing that stood out to me is that one of the playlists contains a whole album, which is not like the other playlists.

## Column 2 {data-width = 350}

### Volgende


# Visualisation

## Column 1

### First graph

For the first plot, I put danceability and energy on the axes. I thought these were two good features to see the difference between the parts of the day.

```{r}
lunch <- get_playlist_audio_features("","15XTzdCk80I6KrHwrRgR0L")
rock_fissa <- get_playlist_audio_features("", "7fLp6h0KxTCnNWsbBpSWXi")
bar_playlists <-
  bind_rows(
    lunch %>% mutate(category = "Lunch"),
    rock_fissa %>% mutate(category = "Avond"))

plot_1 <- bar_playlists %>% mutate(
    mode = ifelse(mode == 0, "Minor", "Major")
  ) %>%ggplot(                     
    aes(
      x = valence,
      y = energy,
      size = loudness,
      colour = mode
    )
  ) +
  geom_point() +              
  geom_rug(size = 0.1) +      
  geom_text(                  
    aes(
      x = valence,
      y = energy,
      label = label
    ),
    data = 
      tibble(
        label = c("", ""),
        category = c("Avond", "Lunch"),
        valence = c(0.090, 0.123),
        energy = c(0.101, 0.967)
      ),
    colour = "black",        
    size = 3,                
    hjust = "left",          
    vjust = "bottom",        
    nudge_x = -0.05,          
    nudge_y = 0.02            
  ) +
  facet_wrap(~category) +     
  scale_x_continuous(         
    limits = c(0, 1),
    breaks = c(0, 0.50, 1),   
    minor_breaks = NULL       
  ) +
  scale_y_continuous(         
    limits = c(0, 1),
    breaks = c(0, 0.50, 1),
    minor_breaks = NULL
  ) +
  scale_colour_brewer(        
    type = "qual",            
    palette = "Paired"        
  ) +
  scale_size_continuous(      
    trans = "exp",            
    guide = "none"            
  ) +
  theme_light() +             
  labs(                       
    x = "Danceability",
    y = "Energy",
    colour = "Mode"
  )

ggplotly(plot_1)
```

# Conclusion


